### 2024-02-19 revision/reply_to_reviewer1.tex

%	NICK ADD PLOT
%	
%	In particular, we note the poor performance for the Ricker wavelet-type kernel (right two columns of Figure 7). We observed a similar configuration for the impulse responses for the Hessian and Gauss-Newton Hessian in wave inverse problems.

%	 The configuration of negative numbers in this example is similar The negative numbers in the integral kernel for the wave Hessian (and Gauss-Newton Hessian) are distributed in a manner similar to the Ricker wavelet-type impulse response shown in the right two columns of Figure 7.
%	
%	 We also added a discussion of the impact of non-locality and negativity on the performance of the method (Section 3.2).
%	
%	The 
%	
%	(1) Three-figure plot: Stokes, Advection/diffusion, wave impulse responses. 
%	
%(2) showing that the method performs poorly for the ricker example
%	
%Nick will try to get an impulse response for the wave
%  problem.
}

\item \referee{It is interesting to see the effectiveness of the
  approximate Hessian used as a preconditioner. It would also be very
  interesting to see how accurate the approximate Hessian is compared
  to a full Hessian in terms of the number of batches, ellipsoid
  sizes, and in particular the total number of impulse responses. Can
  you say anything about the convergence of the approximation, either
  numerical or theoretical?}

  \noindent{
  	In Figure 8 in the revised manuscript we added a convergence study for the blur kernel example. We show the convergence of the proposed PSF method as a function of the total number of impulse responses and the number of batches, for several ellipsoid size parameters $\tau$ ranging from $2.0$--$4.0$. We observe a roughly linear convergence rate until a limit, which depends on $\tau$, is reached. The larger $\tau$ is, the smaller the limit is on the achievable error. Before this limit, the error decreases at roughly the same rate for all $\tau$ in terms of the number of impulse responses. This means that the error decreases faster for smaller $\tau$ as a function of the number of batches, since smaller $\tau$ means more impulse responses per batch. Basically, taking $\tau$ larger makes the method converge more slowly, but allows it to eventually achieve a lower level of error. We added a discussion of this convergence study in Section 7.1 of the revised manuscript.

%  	 I.e., the relative error decreases as $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$ or $\text{constant} \times \# \text{batches}^{-1}$ until the limit is reached, then the error does not decrease any further. Increasing $\tau$ lowers this limit, allowing the method to achieve higher accuracy. Before this limit is reached, the convergence rate is the same for all $\tau$ as a function the number of impulse responses.

%  	Figure 6 illustrates how the error decreases with increasing numbers of batches and therefore impulse responses. ALSO TABLE 4 CONDITION NUMBER
%  	
%  	Frog example with a=1. Error vs \#batches, for different tau. Also do a plot with \#impulses on the x-axis.
%  	
%  	Nick will try to create a plot that shows some error
%    (e.g. Frob norm, cond number) as a function of number of impulse
%    responses.
}

### 2024-02-19 revision/reply_to_reviewer1.tex

\paragraph{General Comments:}

\referee{I recommend to accept the paper for publication with minor
  revisions. I suggest the authors properly address the following
  questions to help for its better understanding and broader
  applications.\\}

\begin{enumerate} 
\item \referee{The author mentioned one limitation of the proposed method:
      the Hessian should have a local non-negative integral kernel,
      which is not satisfied for wave inverse problems that lead to a
      substantial amount of negative entries of the Hessian. It would
      be interesting if the author could elaborate more details with
      intuition or numerical evidence for this property and the
      limitation, especially if you also use Gauss-Newton approximate
      Hessian for the wave inverse problems as in this paper.}

\noindent{
	We added a numerical study investigating the robustness of the method to violations of the non-negative kernel assumption on two two-dimensional blurring kernel examples (Figure 7). We also added a discussion of the impact of non-locality and negativity on the performance of the method (Section 3.2).

	One takeaway message from Figure 7 is that the impact of negative numbers depends strongly on the spatial configuration of these negative numbers---in the first example (columns~1 and~2 in Figure~7), the negative numbers have minimal impact on the ellipsoid impulse response support estimates, whereas in the second example (columns~3 and~4), even a small amount of negative numbers causes the impulse response support estimate to be poor. The configuration of negative numbers for the integral kernel in wave inverse problem Hessians (including Gauss-Newton Hessians) is similar to that of the second example, and therefore our method is unlikely to be effective for wave Hessians.

@@

	In Figure \ref{fig:marmousi_impulse} here (but not in the paper due to page limitations), we show an impulse response for the Gauss-Newton Hessian for the Marmousi wave inverse problem. Note that the non-local component of the impulse response shifts the mean far away from the visual ``center'' of the impulse response, and causes the ellipsoid impulse response support estimate to be effectively the entire domain.

	We also added more references and discussion regarding the wave inverse problem Hessian at the end of Section 2 in the revised manuscript.

	\begin{figure}
		\centering
		\includegraphics[scale=0.735]{marmousi_impulse_response_full_domain_ellipsoid_with_tau_annotations.png}
		\includegraphics[scale=0.3975]{marmousi_impulse_response_zoomed_in.png}

		\caption{Impulse response for the Gauss-Newton Hessian for the Marmousi wave inverse problem. On the left we show the impulse response on the whole domain. The color scale is truncated for better visualization of the non-local transmission components. The estimated mean of the impulse response is shown as a red dot. Estimated support ellipsoids with $\tau\in \{1.0, 2.0, 3.0\}$  are shown as black ellipses. On the right we show a zoomed in impulse response, showcasing the negative numbers in the integral kernel. The data for this impulse response was kindly provided to us by Mathew Hu (UT Austin).
		}
		\label{fig:marmousi_impulse}
	\end{figure}

@@

%	NICK ADD PLOT
%	
%	In particular, we note the poor performance for the Ricker wavelet-type kernel (right two columns of Figure 7). We observed a similar configuration for the impulse responses for the Hessian and Gauss-Newton Hessian in wave inverse problems.

%	 The configuration of negative numbers in this example is similar The negative numbers in the integral kernel for the wave Hessian (and Gauss-Newton Hessian) are distributed in a manner similar to the Ricker wavelet-type impulse response shown in the right two columns of Figure 7.
%	
%	 We also added a discussion of the impact of non-locality and negativity on the performance of the method (Section 3.2).
%	
%	The 
%	
%	(1) Three-figure plot: Stokes, Advection/diffusion, wave impulse responses. 
%	
%(2) showing that the method performs poorly for the ricker example
%	
%Nick will try to get an impulse response for the wave
%  problem.
}

\item \referee{It is interesting to see the effectiveness of the
  approximate Hessian used as a preconditioner. It would also be very
  interesting to see how accurate the approximate Hessian is compared
  to a full Hessian in terms of the number of batches, ellipsoid
  sizes, and in particular the total number of impulse responses. Can
  you say anything about the convergence of the approximation, either
  numerical or theoretical?}

  \noindent{
  	In Figure 8 in the revised manuscript we added a convergence study for one of the blur kernel examples. We show the convergence of the proposed PSF method as a function of the total number of impulse responses and the number of batches, for several ellipsoid size parameters $\tau$ ranging from $2.0$--$4.0$. We observe a roughly linear convergence rate until a limit, which depends on $\tau$, is reached. The larger $\tau$ is, the smaller the limit is on the achievable error. Before this limit, the error decreases at roughly the same rate for all $\tau$ in terms of the number of impulse responses. This means that the error decreases faster for smaller $\tau$ as a function of the number of batches, since smaller $\tau$ means more impulse responses per batch. Basically, with larger $\tau$ the method converge more slowly than with smaller $\tau$, but larger $\tau$ allows the method to eventually achieve a lower level of error. We added a discussion of this convergence study in Section 7.1 of the revised manuscript.

  	Although the method can achieve high accuracy, we would like to emphasize that it is targeted at achieving modest accuracies (e.g., $5\%$ error) to build preconditioners for problems that are challenging for existing methods. In Table 2 in the revised manuscript, following the other reviewer's suggestion, we added a comparison of the computational cost
  	(measured in operator applies) to approximate the kernel in one of the spatially
  	varying blurring kernel examples. We compare the results using the proposed PSF method,
  	the randomized HODLR (hierarchical off diagonal low rank) method,
  	and GLR (global low rank) approximation using randomized SVD. The
  	results show an overwhelming computational gain.

@@

%  	 I.e., the relative error decreases as $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$ or $\text{constant} \times \# \text{batches}^{-1}$ until the limit is reached, then the error does not decrease any further. Increasing $\tau$ lowers this limit, allowing the method to achieve higher accuracy. Before this limit is reached, the convergence rate is the same for all $\tau$ as a function the number of impulse responses.

%  	Figure 6 illustrates how the error decreases with increasing numbers of batches and therefore impulse responses. ALSO TABLE 4 CONDITION NUMBER
%  	
%  	Frog example with a=1. Error vs \#batches, for different tau. Also do a plot with \#impulses on the x-axis.
%  	
%  	Nick will try to create a plot that shows some error
%    (e.g. Frob norm, cond number) as a function of number of impulse
%    responses.
}

 \item \referee{The paper presented a nice complexity analysis of the
   steps in constructing the approximate Hessian, with the dominating
   cost arising from the Hessian matrix-vector product. It would be
   interesting to see the computational time of each step for the
   numerical examples to provide a good sense of the complexity and to
   support the 5--10X computational reduction by accounting also the
   overhead beyond the PDE solves.
}

   \noindent{This is a fair point. We agree with the reviewer that there is an overhead
     beyond the PDE solves. However, since we have not optimized our codes for performance, it would be
     perhaps misleading to report profiling/timing information. Nevertheless, for all numerical tests we
     present in this paper, the cost to build the H-matrices (item (2) in Section 6) was small compared to
     PDE solves (item (1) in Section 6). The further H-matrix linear algebra operations (item (3) in Section 6), as implemented in the HLIBPro library, were less costly than the PDE solves, but their costs were not negligible. In particular, enforcing positive definiteness of the H-matrix via the procedure presented in Section 5.5 was costly. Because of the polylog linear complexity of H-matrix operations, for large-scale problems these H-matrix overhead costs would eventually become insignificant compared to the PDE solves.
%     
%     However, for all numerical tests we
%     present in this paper, this cost was small compared to
%     PDE-applies. This is due to the need for a relatively small
%     number of sample points ($m$) in all batches. Since we have not
%     invested time to optimize our codes for performance, it would be
%     perhaps misleading to report profiling-timing information. When
%     $m$ is very large, one can reduce this cost with more involved
%     computational geometry methods. This was beyond the scope of our
%     work.
}

### 2024-02-19 paper_sisc_first_revision/localpsf.tex
respond to reviewer 1 question 2
edits during discussion between Nick and Noemi

We approximate the support of $\impulseresponse_x$ with the ellipsoid
%
%
\begin{equation}
\label{eq:support_ellipsoid}
E_x := \{x' \in \Omega: (x' - \spatialmean(x))^T \spatialcov(x)^{-1} (x' - \spatialmean(x)) \le \tau^2\},
\end{equation}
%
%
where $\tau$ is a fixed constant (see Figure \ref{fig:frog_moments_ellipsoid}). The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\impulseresponse_x$. The quantity $\tau$ is a parameter that must be chosen appropriately. The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$. However, in Section~\ref{sec:sample_point_selection} we will see that the cost of the proposed method depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids is better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$. In practice, we find that $\tau=3.0$ yields a reasonable balance between these competing interests, and use $\tau=3.0$ in the numerical results. The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is typically conservative.

@@

@@

@@

\begin{table}
	\begin{center}
	{\small
	\begingroup
	\setlength{\tabcolsep}{8pt}
	\renewcommand{\arraystretch}{1.0}
	\begin{tabular}{c| c | c c c c c}
				& Error tol.    &	\#applies PSF   & \#applies HODLR 	& \#applies RSVD 	\\%& rank \\ 
	\hline 
				& 20\%    		& 11  				& 592				& 354				\\%& 111  \\ 
	$L=1$ 		& 10\% 			& 15  				& 772				& 520				\\%& 178  \\ 
				& 5\% 			& 23   				& 924				& 674				\\%& 244 \\ 
	\hline
				& 20\%    		& 8  				& 852 				& 1316				\\%& 417 \\ 
	$L=1/2$ 	& 10\%    		& 10  				& 1144				& 1916				\\%& 665 \\ 
				& 5\% 	  		& 12   				& 1404				& 2456				\\%& 903 \\ 
	\hline
				& 20\%    	 	& 7					& 932				& 2624				\\%& 845  \\ 
	$L=1/3$ 	& 10\%    		& 8  				& 1264 				& 3734				\\%& 1321  \\ 
				& 5\% 			& 8   				& 1520 				& 4660				\\%& 1759  \\ 
	\end{tabular}
	\endgroup
	}
	\end{center}
	\caption{(Spatially variant blur) Comparison of cost to
          approximate the blur kernel from Equation
          \eqref{eq:frog_kernel} using the PSF method, the randomized
          HODLR (hierarchical off diagonal low rank) method, and GLR
          (global low rank) approximation using randomized SVD. The
          quantity $L$ scales the width of the impulse responses,
          hence it influences the rank of the operator. Large $L$
          means low rank, and small $L$ means high rank. The second
          column (``Error tol'') is the relative error in the
          approximation of the kernel measured in the Frobenius norm,
          $||\boldsymbol{\Phi} -
          \widetilde{\boldsymbol{\Phi}}||_\text{Fro} /
          ||\boldsymbol{\Phi}||_\text{Fro}$. The remaining three
          columns show the number of operator applies required to
          achieve the given error tolerances, using the PSF, HODLR,
          and GLR methods.}
	\label{tab:frog_psf_vs_hodlr_vs_glr}
\end{table}

\begin{figure}
	\begin{subfigure}{0.52\textwidth}
		\begin{center}
			\includegraphics[scale=1.0]{frog_tau_impulses.pdf}
		\end{center}
	\end{subfigure}
	\begin{subfigure}{0.47\textwidth}
		\begin{center}
			\includegraphics[scale=1.0]{frog_tau_batches.pdf}
		\end{center}
	\end{subfigure}
	\caption{Relative error in the PSF approximation of the kernel, $||\boldsymbol{\Phi} - \widetilde{\boldsymbol{\Phi}}||_\text{Fro}/||\boldsymbol{\Phi}||_\text{Fro}$, for different ellipsoid size parameters, $\tau$. Left: Relative error vs. the total number of impulse responses used in the approximation. Right: Relative error vs. the number of impulse response batches. The dashed gray lines show linear convergence rates, i.e., $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$ on the left, and  $\text{constant} \times \# \text{batches}^{-1}$ on the right.}
	\label{fig:PSF_tau_convergence}
\end{figure}

>>>>

We approximate the support of $\impulseresponse_x$ with the ellipsoid
%
%
\begin{equation}
\label{eq:support_ellipsoid}
E_x := \{x' \in \Omega: (x' - \spatialmean(x))^T \spatialcov(x)^{-1} (x' - \spatialmean(x)) \le \tau^2\},
\end{equation}
%
%
where $\tau$ is a fixed constant (see Figure \ref{fig:frog_moments_ellipsoid}). The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\impulseresponse_x$. The quantity $\tau$ is a parameter that must be chosen appropriately. The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$. However, in Section~\ref{sec:sample_point_selection} we will see that the cost of the proposed method depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids is better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$. In practice, we find that $\tau=3.0$ yields a reasonable balance between these competing interests, and use $\tau=3.0$ in all numerical results, except for Figure \ref{fig:PSF_tau_convergence}, where we study the effects of varying $\tau$. The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is typically conservative.

@@

\begin{table}
	\begin{center}
		{\small
			\begingroup
			\setlength{\tabcolsep}{8pt}
			\renewcommand{\arraystretch}{1.0}
			\begin{tabular}{c| c | c c c c c}
				& Error tol.    &	\#applies PSF   & \#applies HODLR 	& \#applies RSVD 	\\%& rank \\ 
				\hline 
				& 20\%    		& 11  				& 592				& 354				\\%& 111  \\ 
				$L=1$ 		& 10\% 			& 15  				& 772				& 520				\\%& 178  \\ 
				& 5\% 			& 23   				& 924				& 674				\\%& 244 \\ 
				\hline
				& 20\%    		& 8  				& 852 				& 1316				\\%& 417 \\ 
				$L=1/2$ 	& 10\%    		& 10  				& 1144				& 1916				\\%& 665 \\ 
				& 5\% 	  		& 12   				& 1404				& 2456				\\%& 903 \\ 
				\hline
				& 20\%    	 	& 7					& 932				& 2624				\\%& 845  \\ 
				$L=1/3$ 	& 10\%    		& 8  				& 1264 				& 3734				\\%& 1321  \\ 
				& 5\% 			& 8   				& 1520 				& 4660				\\%& 1759  \\ 
			\end{tabular}
			\endgroup
		}
	\end{center}
	\caption{(Blur) Comparison of cost to
		approximate the blur kernel from Equation
		\eqref{eq:frog_kernel} using the PSF method, the randomized
		HODLR (hierarchical off diagonal low rank) method, and GLR
		(global low rank) approximation using randomized SVD. The
		quantity $L$ scales the width of the impulse responses,
		hence it influences the rank of the operator. Large $L$
		means low rank, and small $L$ means high rank. The second
		column (``Error tol'') is the relative error in the
		approximation of the kernel measured in the Frobenius norm,
		$||\boldsymbol{\Phi} -
		\widetilde{\boldsymbol{\Phi}}||_\text{Fro} /
		||\boldsymbol{\Phi}||_\text{Fro}$. The remaining three
		columns show the number of operator applies required to
		achieve the given error tolerances, using the PSF, HODLR,
		and GLR methods.}
	\label{tab:frog_psf_vs_hodlr_vs_glr}
\end{table}

@@

\begin{figure}
	\begin{subfigure}{0.52\textwidth}
		\begin{center}
			\includegraphics[scale=1.0]{frog_tau_impulses.pdf}
		\end{center}
	\end{subfigure}
	\begin{subfigure}{0.47\textwidth}
		\begin{center}
			\includegraphics[scale=1.0]{frog_tau_batches.pdf}
		\end{center}
	\end{subfigure}
	\caption{(Blur) Relative error 
		%		in the PSF approximation of the kernel
		%	$||\boldsymbol{\Phi} - \widetilde{\boldsymbol{\Phi}}||_\text{Fro}/||\boldsymbol{\Phi}||_\text{Fro}$,
		for different ellipsoid size parameters, $\tau$, vs. the total number of impulse responses (left), and the number of batches (right). 
		%	Left: Relative error vs. the total number of impulse responses used in the approximation. Right: Relative error vs. the number of impulse response batches. 
		The dashed gray lines show linear convergence rates, i.e., $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$ on the left, and  $\text{constant} \times \# \text{batches}^{-1}$ on the right.}
	\label{fig:PSF_tau_convergence}
\end{figure}

In Figure \ref{fig:PSF_tau_convergence} we show the convergence of the proposed PSF method on the blur kernel as a function of the total number of impulse responses (left), and the number of batches (right). We show convergence for several ellipsoid size parameters $\tau$, ranging from $2.0$--$4.0$. The results in Figure \ref{fig:PSF_tau_convergence} (left) show that the relative error decreases as $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$
% or $\text{constant} \times \# \text{batches}^{-1}$, 
suggesting linear convergence. The linear convergence stalls at a limit which depends on $\tau$. Increasing $\tau$ lowers this limit, allowing the method to achieve higher accuracy. 
In Figure \ref{fig:PSF_tau_convergence} (right), the results show that before this limit is reached, the convergence is faster for smaller $\tau$ in terms of the number of batches. This is expected because smaller $\tau$ results in more impulse responses per batch. Larger $\tau$ causes the method converge more slowly than smaller $\tau$, but with larger $\tau$ the method stalls at a lower level of error than it does with smaller $\tau$ (see, e.g., $\tau=4.0$ vs. $\tau=2.5$).

%----
%, but allows the method to eventually achieve a lower level of error.
%
%----
%
%For smaller $\tau$ regime, we observe faster convergence
%
%Before this limit is reached, the convergence rate is the same for all $\tau$ as a function the number of impulse responses. This means that the convergence is faster for smaller $\tau$ as a function of the number of batches, since smaller $\tau$ means more impulse responses per batch. In summary, taking $\tau$ larger makes the method converge more slowly, but allows it to eventually achieve a lower level of error.

@@
